
\section{Matrix Algebras}

\subsection{Scaled matrix product}

Let $\M_n$ be the set of $n \times n$ matrices with complex entries equipped
with a multiplication:
\begin{eqnarray*}
  m_y :& \M_n \otimes \M_n &\to     \M_n \\
      :&  a   \otimes b    &\mapsto yab
\end{eqnarray*}
for some non-zero $y \in \C$. It is easy to see that this is an associative,
bilinear multiplication on $\M_n$ with unit:
\begin{eqnarray*}
  e_y :& \C &\to     \M_n \\
    :&  1 &\mapsto y^{-1}I_n
\end{eqnarray*}
where $I_n$ is the $n \times n$ identity matrix. We define
$w_y : = m_y^{\dagger}$ and $c _y := e_y^{\dagger}$.

When $y = 1$, $m_y = m_1$ is the usual matrix multiplication with the usual
unit. In this case, $w_1$ is given by the following \cite[8]{CatQChan}:
\[
  A = \sum_{i, j} a_{ij}e_{ij}
    \mapsto \sum_{i, j} a_{ij} \sum_{k} e_{ik} \otimes e_{kj}
\]
and $c_1$ is the trace map:
\begin{eqnarray*}
  c_1 :& \M_n &\to     \C \\
      :& \sum_{i, j} a_{ij}e_{ij} &\mapsto \sum_{i} a_{ii}
\end{eqnarray*}

Let the matrices representing $m_1$ and $e_1$, seen as linear maps, be $P_1$ and
$Q_1$ respectively. It is easy to see that the matrices for $m_y$ and $e_y$ are
$P_y := yP_1$ and $Q_y = y^{-1}Q_1$ respectively, so that $m_y = ym_1$ and
$e_y = y^{-1}e_1$, for any $y \in \C$. Hence,
\begin{align*}
  w_y &= m_y^{\dagger} = \cnj{y}m_1^{\dagger} = \cnj{y}w_1 \\
  c_y &= e_y^{\dagger} = (\cnj{y})^{-1}e_1^{\dagger} = (\cnj{y})^{-1}c_1
\end{align*}

It is well known that $(m_1, e_1, w_1, c_1)$ is a normalizable dagger Frobenius
structure on $\M_n$ \cite[8, 10]{CatQChan}. Using a Frobenius law for $m_1$
and $w_1$, we observe that
\[
  (m_y \otimes \id) \circ (\id \otimes w_y)
  = (ym_1 \otimes \id) \circ (\id \otimes (\cnj{y})^{-1}w_1)
  = y(\cnj{y})^{-1}w_1m_1
  = w_ym_y
\]
which establishes a Frobenius law for $m_y$ and $w_y$. We can prove the other
Frobenius law for $m_y, w_y$ similarly, establishing the following theorem.

\begin{thm}
$\M_n$ along with $m_y, e_y, w_y, c_y$ is a dagger Frobenius algebra.
\end{thm}

By a diagram deformation applied on the definition of normalizability
\cite[9]{CatQChan}, the above algebra is normalizable if and only if there
exists a central positive definite map $\fn{z}{\M_n}{\M_n}$ satisfying:
\[
  c_ym_yw_yz^2 = c_y \iff |y|^2c_1m_1w_1z^2 = c_1
\]
We can show that $m_1w_1(a) = na$ and hence $m_yw_y(a) = y\cnj{y} (na)
= n|y|^2 a$ for all $a \in \M_n$ so that the above relation becomes:
\begin{equation}\label{simplenorm}
  c_1z^2 = \frac{1}{n|y|^2}c_1
\end{equation}
An obvious candidate for $z$ is the map $a \mapsto \frac{1}{\sqrt{n}|y|} a$.
However, we are interested in $z$ arising from a PFT. So, we define:
\begin{defn}
Let $(A, m, e, w, c)$ be a dagger Frobenius algebra defined by (or defining) a
PFT $F$. We say that $A$ is normalizable under $F$ if there exists a planar
cobordism $\fn{Z}{I}{I}$ for which $F(Z)$ is a normalizer for $A$.
\end{defn}

Let $F$ be a PFT defined by the Frobenius structure $\M_n, m_y, e_y, w_y, c_y$.
That is, if $I$, $M$, $E$, $W$, $C$ are the interval, pair-of-pants, cup (unit
for $I$ seen as a Frobenius monoid under $M$ in $2\Thick$), co-pair-of-pants and
cap (counit for $I$) respectively, then we have:
\[
  F(I) = \M_n, F(M) = m_y, F(E) = e_y, F(W) = w_y, F(C) = c_y
\]

Let $\M_{n}$ with the above structure be normalizable under $F$, with normalizer
$z = F(Z)$ for some planar cobordism $\fn{Z}{I}{I}$. Noting that every
$\fn{Z}{I}{I}$ is determined by its genus, say $k$, and that each hole
decomposes as the map $MW$, we must have that $Z = (MW)^k$. This yields
\[
  z(a) = (m_yw_y)^k(a) = n^{k}|y|^{2k}a
\]
for all $a \in \M_n$. Substituting this into the simplified normalizability
condition \eqref{simplenorm}, we get:
\begin{align*}
       & c_1(z^2(a)) = \frac{1}{n|y|^2} c_1(a)\\
  \iff & n^{2k}|y|^{4k} c_1(a) = \frac{1}{n|y|^2} c_1(a)\\
  \iff & |y|^{4k + 2}c_1(a) = \frac{1}{n^{2k + 1}} c_1(a)
\end{align*}
Since $a$ is arbitrary, we can choose $a = e_{11}$ so that the above relation
becomes:
\[
  |y|^{2(2k + 1)} = \frac{1}{n^{2k + 1}}
  \iff |y| = \frac{1}{\sqrt{n}}
\]

We note that if this is true, then it it so, regardless of the genus of $Z$ --
this is unsurprising because:
\[
  z(a) = n^k|y|^{2k} a = n^k \br{n^{-1/2}}^{2k} a = a
\]
so that $z$ must be the identity. We collect these results as follows.
\begin{thm}\label{pftnorm}
The dagger Frobenius algebra $(\M_n, m_y, e_y, w_y, c_y)$ is normalizable under
a PFT if and only if $|y| = n^{-1/2}$. In this case, $F(Z)$ is a normalizer for
every planar cobordism $\fn{Z}{I}{I}$ in $2\Thick$ and must be be the identity
map $\M_n \to \M_n$.
\end{thm}
\begin{proof}
If $z = F(Z)$ for some PFT $F$ and some planar cobordism $Z$ in $2\Thick$, then
we have shown above that $|y| = n^{-1/2}$. If, on the other hand,
$|y| = n^{-1/2}$, then we can set $Z = (MW)^k$ for any $k \in \N \cup \set{0}$
and $z = F(Z)$. In this case, using the decomposition of a hole as $MW$, we
get:
\[
  c_1(z^2(a)) = c_1(n^k|y|^{2k} a) = n^k \frac{1}{n^k} c_1(a) = c_1(a)
    = \frac{1}{n \frac{1}{(\sqrt{n})^2}} c_1(a)
    = \frac{1}{n|y|^2} c_1(a)
\]
Thus, $F(Z)$ is a normalizer for the given dagger Frobenius structure. That
$F(Z)$ must be the identity was shown before.
\end{proof}

\begin{cor}
Not all normalizable dagger Frobenius algebras are accounted for by PFTs.
\end{cor}
\begin{proof}
$(\M_n, m_1, e_1, w_1, c_1)$ is a normalizable dagger Frobenius algebra with
normalizer $a \mapsto n^{-1/2} a$ but this normalizer cannot be not in the
image of a PFT $F$.
\end{proof}

We do not wish to give up just yet. We let $\M_{n, y}$ denote the dagger
Frobenius algebra $(\M_n, m_y, e_y, w_y, c_y)$ and consider the map
$\phi : \M_{n, y} \to \M_{n, 1} : a \mapsto \alpha a$, for some non-zero
$y, \alpha \in \C$ and observe that
\[
  \phi m_y(a \otimes b) = \phi(yab) = y\phi(ab) = y\alpha ab
\]
and
\[
  m_1(\phi(a) \otimes \phi(b)) = \alpha^2 ab
\]
For phi to be an algebra homomorphism, we need $\alpha^2 = y\alpha \iff y =
\alpha$. Indeed, we see that, under phi,
\[
  y^{-1}I_n \mapsto I_n
\]
showing unitality. Invertibility being clear, $\phi$ is an algebra isomorphism.
We observe the interaction of $\phi$ with the dagger structure:
\[
  \phi(a^{\dagger}) = ya^{\dagger} = (\cnj{y}a)^{\dagger} =
    \frac{y}{\cnj{y}}\phi(a)^{\dagger}
\]
If $y$ is real, then $\phi$ is a $*$--isomophism.
\begin{thm}
Every $\M_{n, 1}$ is $*$--isomorphic to $\M_{n, y}$ for some real $y$.
\end{thm}

If $\M_{n, y}$ is normalizable under a PFT, then for every $\fn{Z}{I}{I}$,
$F(Z) = \id_{\M_{n, y}}$ is a normalizer for $\M_{n, y}$ and
$\phi \circ F(Z) = \phi$. Furthermore, $|y| = n^{-1/2}$ by \ref{pftnorm}.
Finally, if $y$ is real, making $\phi$ a $*$--isomorphism as above, we must have
$y = \pm n^{-1/2}$. Picking the positive value for $y$,
$\phi \circ F(Z) = \phi$ seen as map $\M_{n, 1} \to \M_{n, 1}$ is the usual
normalizer for $\M_{n, 1}$.

I am guessing that this argument generalizes to direct sums of the form
$\bigoplus_{i = 1}^r \M_{n_i}$ as in \cite[10]{CatQChan}.  If this is true, we
may have characterized the normalizability of finite dimensional $C^*$--algebras
through PFTs up to $*$--isomorphism, since normalizable dagger Frobenius
algebras over $\C$ are precisely the finite direct sums of finite dimensional
matrix algebras (up to $*$--isomorphism?). In other words, if the above is
correct, then
\begin{thm}
Every finite dimensional $C^*$--algebra is $*$--isomorphic to a dagger Frobenius
algebra normalizable under a PFT with the ismorphism ``carrying over the
normalizer''\footnote{in some sense} both ways.
\end{thm}

If we are satisfied with the above, we may also look at the $\text{CP}^*$
condition or complete-positivity of mappings between $C^*$ algebras and their
relations with monoidal natural transformations between PFTs.

\subsection{Matrix product with intervening matrix}

We now think of more interesting multiplications. For instance, we can take the
map
\[
  m_r : (a \otimes b) \mapsto arb
\]
for some invertible $r \in \M_n$. This is easliy seen to be associative:
\[
  (arb)rd = ar(brd)
\]
with unit $e_r : 1 \mapsto r^{-1}$:
\[
  arr^{-1} = a = r^{-1}ra
\]
This product also inherits bilinearity from matrix multiplication:
\[
  (p_1a_1 + p_2a_2)r(p_3a_3 + p_4a_4)
    = p_1p_3a_1ra_3 + p_1p_4a_1ra_4 + p_2p_3a_2ra_3 + p_2p_4a_2ra_4
\]

We define a useful map:
\begin{defn}
We call the map $\fn{u_n}{\M_n}{\C^{n^2}}$ defined by
\[
  u_n\br{\bmat{a_{11} & \dots & a_{1n} \\
             \vdots & \ddots & \vdots\\
             a_{n1} & \dots & a_{nn}}}
    = \bmat{a_{11} & \dots & a_{1n} & a_{21} & \dots & a_{2n} &
            \dots & a_{n1} & \dots & a_{nn}}^{T}
\]
a row serialization\footnote{This is a computer science term, in case the reader
is wondering -- serialization typically refers to the representation of any kind
of data as an array of numbers.} of $\M_n$ and the map $v_n := u_n((-)^T)$, a
column serialization of $\M_n$.
\end{defn}

\begin{rmk}
It might be convenient to note that we are choosing the convention of sending
matrices to column vectors as opposed to row vectors. At the same time, $u_n(a)$
and $v_n(a)$ are not transposes of each other -- they are both column vectors
but have different orderings of the entries of the matrices.
\end{rmk}

Noticing that $u_n$ and $v_n$ are isomorphisms, we can define maps
$\M_n \to \M_m$ by defining maps $\C^{n^2} \to \C^{m^2}$, given by
$m^2 \times n^2$ matrices. Viewing $\M_n$ as $\C^{n^2}$ and $\M_n \otimes \M_n$
as $\C^{n^4}$ under $u_n$ and $u_{n^2}$ respectively, let $M_r$ be the
$n^2 \times n^4$ matrix representing
\[
  \fn{m_r}{\M_n \otimes \M_n \cong \C^{n^4}}{\C^{n^2} \cong \M_n}
\]
where we are taking $\otimes$ as the Kronecker. We can then compute
$u_{n^2}(a \otimes b)$ and try to get $arb$ from this product, by applying a
linear map on it. If we let
\[
  a = [a_{ij}], b = [b_{ij}], r = [r_{ij}]
\]
then it is straightforward to check that $M_r$ is the following
matrix\footnote{In reality, I only checked this for $n = 1, \dots, 5$ and
concluded the same for all $n$, because that is how logic works!}:
\[
  \begin{bmatrix}
    v_n(r)^T  & 0_n       & \dots   & 0_n     \\
    0_n       & v_n(r)^T  & \dots   & 0_n     \\
    \vdots    & \vdots    & \ddots  & \vdots  \\
    0_n       & 0_n       & \dots   & v_n(r)^T
  \end{bmatrix} \otimes I_{n}
\]
where the left multiplicand has $n$ rows, $I_n$ is the $n \times n$ identity
matrix and $0_n$ is the row vector consisting of $n$ $0$'s. Let
$W_r$ be the conjugate transpose of $M_r$ determining the adjoint $w_r$ of
$m_r$. We also define $c_r$ as the conjugate transpose of $e_r$. I believe that
since $(-)^{\dagger}$ is a contravariant functor, it automatically makes
$(w_r, c_r)$ a coalgebra structure on $\M_n$.\footnote{Confirmation of this
would be great.}

We can use the matrix for $M_r$ to show that
\begin{align*}
  (m_r \otimes \id) \circ (\id \otimes w_r) = w_rm_r
  &\iff (M_r \otimes I_{n^2}) \cdot (I_{n^2} \otimes W_r) = W_r \cdot M_r\\
  (\id \otimes m_r) \circ (w_r \otimes \id) = w_rm_r
  &\iff (I_{n^2} \otimes M_r) \cdot (W_r \otimes I_{n^2}) = W_r \cdot M_r
\end{align*}
making $(\M_n, m_r, e_r, w_r, c_r)$ a dagger Frobenius algebra.\footnote{Again,
I assumed that showing these for small $n$ shows them for all $n$!}

It is also straightforward to check that
\[
  M_rW_r = M_rM_r^{\dagger}
    = \br{\sum_{i, j} |r_{ij}|^2}I_{n^2}
\]
Denoting $||r|| := \br{\sum_{i, j} |r_{ij}|^2}^{\frac{1}{2}}$, this shows that
$m_rw_r(a) = ||r||^2a$ for all $a \in \M_n$. We now wish to compute
$c_r(m_rw_r)^k$. For this, we need to compute $c_r := e_r^{\dagger}$. Now,
$e_r$ is given by the matrix
\[
  E_r := u_n(r^{-1})
\]
so that $c_r$ is given by the matrix:
\[
  C_r = u_n(r^{-1})^{\dagger}
\]
Note that $C_r$ is the conjugate transpose of $u_n(r^{-1})$, not that of
$r^{-1}$ itself so that to compute $C_r$ we take only the conjugate of
$r^{-1}$ and treat it as a row vector as opposed to a column vector (after row
serialization, of course).

Letting $r^{-1} = [r'_{ij}]$, we have
\[
  c_r(a) = C_ru_n(a) = \sum_{i, j} \cnj{r'_{ij}}a_{ij}
\]
and
\[
  c_r(m_rw_r)^k(a) = ||r||^{2k} \sum_{i, j} \cnj{r'_{ij}}a_{ij}
\]
Finding a manageable formula for $r'_{ij}$ or the $(i, j)$--entry of $r^{-1}$
seems central to understanding the above expression and, by extension, how PFTs
interact with these ``$arb$'' matrix algebras. Doing a few computations on Sage
with the greatest matrix inversion algorithm in the universe
\cite{SteveInvFormula}, I was not sure if a simple formula can be found, even
using this beautiful method.

Another formula exists that tells us that the formula for $r'_{ij}$ might be
complicated if we do not impose some constraints. If we denote the
$(i, j)$--cofactor of $r$ as $D_{ij}(r)$, then it is well known that:
\[
  r^{-1} = \frac{1}{\det(r)}\bmat{
    D_{11}(r) & \dots   & D_{1n}(r) \\
    \vdots    & \ddots  & \vdots \\
    D_{n1}(r) & \dots   & D_{nn}(r)
  }^{T}
\]
so that
\[
  r'_{ij} = \frac{1}{\det(r)}D_{ji}(r)
\]
and
\[
  c_r(m_rw_r)^k(a) = ||r||^{2k} \frac{
      \sum_{i, j} \cnj{D_{ji}}a_{ij}
    }{
      \sum_{p} \cnj{D_{1p}r_{1p}}
    }
\]

I am not sure if there is a particularly simple formula for the right-hand side
for an arbitrary invertible matrix $r$. If there is, then we are in luck --
otherwise, we should, perhaps, look at restrictions on $r$, in addition to
invertibility. At any rate, I think I need some guidance at this point.

\section{Future Work}

We can now try to compute whether this is a Frobenius algebra, but at this
point, we might think of classifying bilinear multiplications on matrix
algebras. When is a morphism $\M_n \otimes \M_n \to \M_n$ associative, unital,
dagger Frobenius and normalizable? I am not sure how easy that is.

Then, we can also think about non-associative or non-unital Frobenius structures
like a coproduct on a Lie algebra that satisfies Frobenius laws, if that is even
possible. A google search with such terms reveals that these terms do exist
together in the literature -- I haven't looked deeper than that.

