
\pagebreak

\thispagestyle{frontstyle}

\section*{Acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}

I am grateful to Dr. Steven Rayan for introducing me to the central idea of this
project, giving me the opportunity to work on it, and for guiding and motivating
me throughout the endeavour. Furthermore, the wonderful courses I took with him
-- Galois Theory and Topology, Algebraic Geometry Through Moduli and Vector
Calculus -- taught me the majority of the geometric background I needed for this
project. At the same time, his counsel and unwavering moral support in various
matters, academic and personal, have made it possible for me to persevere
through trying times over the last two years.

I am indebted to Dr. Cameron Franc and Dr. Christopher Dutchyn for jointly
organizing and teaching a reading course in topos theory which equipped me with
much of the categorical background as well as the initial courage for taking on
this project. This course also made me realize that it is geometry and topology
framed in the language of categories that truly appeal to my mathematical
tastes. I owe much thanks to Dr. Ebrahim Samei for teaching me functional
analysis which made it possible for me to start studying quantum information and
computing, as well as for introducing me to the beautiful subject of geometric
group theory which helped in nurturing my passion and intuition for geomtry and
topology. I am again thankful to Dr. Franc, Dr. Dutchyn, Dr. Samei, Dr. Jenna
Rajchgot and Dr. Christopher Duffy for the fascinating subjects they taught me
as well as for their advice and continued support in all matters.

Finally, my parents and all members of my family -- most importantly, my uncle
Tafiqul Hassan and my aunt Meherun Nahar -- have supported me throughout my
undergraduate life in more ways than I can enumerate. I owe them my deepest
gratitude.

